---
output:
  html_document: default
  pdf_document: default
---

## Course: Visual Analytics for Policy and Management
###Univariate Exercise 1
####Isabella Sun



####Figures:

1. [Ordered Bar Plot using ggplot](#ex1)

2. [Loliplot](#ex2)

3. [Box Plot](#ex3)

4. [Histogram](#ex4)

5. [Histogram](#ex5_1) and [Density Plot using ggplot](#ex5_2)




####Exercise 1 - Ordered Bar Plot
Loading data to be used and libraries:
```{r}
link='https://github.com/EvansDataScience/VisualAnalytics_2_tabularData/raw/master/data/eduwa.rda'

load(file=url(link))

library(ggplot2)

```

Create a table with the proportions and multiply by 100 to get percentages
```{r}
absoluteT=table(eduwa$LocaleType,exclude = 'nothing')
ToPlot=prop.table(absoluteT)
names(ToPlot)[5]='Unknown'
ToPlot=ToPlot*100
paste0(round(ToPlot,2),'%')

```

Create an ordered dataframe for ggplot to plot. "Locale", our categorical variable is not ordered so we can rearange it. 
```{r}

OrderedToPlot=sort(ToPlot) #first ordering the table
tableFreq=as.data.frame(OrderedToPlot) #save the ordered table as a data frame
names(tableFreq)=c("locale","pct") # label vars

tableFreq
```

Misc plot elements like title and source:
```{r}
titleText='Are all locations getting a fair share of public schools in WA?'
sourceText='Source: US Department of Education'
```

<a id='ex1'></a>
Create Plot
```{r}
base= ggplot(data = tableFreq, 
             aes(x = locale,
                 y = pct)) 

plot1 = base + geom_bar(fill ="gray",
                        stat = 'identity') 
plot2 = plot1 + labs(title=titleText,
                     x =NULL, 
                     y = NULL,
                     caption = sourceText)
plot3 = plot2 + geom_hline(yintercept = 25, 
                           linetype="dashed", 
                           size=1.5,
                           alpha=0.5) 

plot4 = plot3 + scale_y_continuous(breaks=c(0,10, 25,40,50),
                                 limits = c(0, 50),
                                 labels=scales::unit_format(suffix = '%')) 

plot5 = plot4 + theme(panel.background = element_rect(fill = "white",
                                                    colour = "grey50"),
                    plot.caption = element_text(hjust = 0), 
                    plot.title = element_text(hjust = 0.5))

plot6 = plot5 + geom_text(aes(y = pct ,
                            label = paste0(round(pct,2), '%')),
                        hjust=1,
                        size = 3)

plot7=plot6 + coord_flip()

plot7
```

####Exercise 2 - Ordered Loliplot
We're using a different dataset for this figure, so let's bring in the new data and load necessary libraries

```{r}
library(foreign)
library(haven)

link='https://github.com/quinhasf/pubpol-599/raw/master/ape_analysis.dta'
chw_fp <- read_dta(url(link)) #download from link and save in datafram named chw_fp

```

Translate information into a dataframe for ggplot 
```{r}
#Create new dataframe with just the relevant variables
LoliToPlot = chw_fp[,c('province','ape_contrib')]

#Order data frame
OrderedLoliToPlot=LoliToPlot[order(LoliToPlot$ape_contrib),]

#Add a new variable and calculate distance from mean
OrderedLoliToPlot$gap=OrderedLoliToPlot$ape_contrib-mean(OrderedLoliToPlot$ape_contrib)

#Another variable indicating whether province is above or below the mean
OrderedLoliToPlot$PositiveGap=ifelse(OrderedLoliToPlot$gap>0,"above average","below average")

```

Misc table elements like title and source 
```{r}
title_loli= "Where is the contribution of community health organizations below average?"
source= "Source: Health Information System"
```

<a id='ex2'></a>
Create Plot
```{r}
base = ggplot(OrderedLoliToPlot, aes(province,gap,color=PositiveGap,
                              label = round(gap,2))) 
lolliplot1 = base + geom_segment(aes(y = 0, 
                                   x = province, 
                                   yend = gap, 
                                   xend = province), color = "gray") 
lolliplot2=lolliplot1 + geom_point() 
lolliplot3= lolliplot2 + scale_x_discrete(limits=OrderedLoliToPlot$province) 

lolliplot5 = lolliplot3 + geom_text(nudge_x=0.45, show.legend = FALSE) 

lolliplot6 = lolliplot5 + labs(title=title_loli,
                     x =NULL, 
                     y = NULL,
                     caption = source)

lolliplot7 = lolliplot6 + theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust=1))


lolliplot7
```

####Exercise 3 - Box Plot
Going back to using the Washington education data...create a data frame with highest grade offered in Washington schools

```{r}
BoxToPlot=as.data.frame(prop.table(table(eduwa$High.Grade)))
names(BoxToPlot)=c('grade','pct')

```

Misc elements like source, title, and labels:
```{r}
boxplot_title = "Distribution of Highest Grade Offered in Washington Schools"
source.text = 'Source: US Department of Education'

#labels
ordLabels= levels(eduwa$High.Grade)
```
<a id='ex3'></a>
Create plot
```{r}
box1 = ggplot(eduwa, aes(y=as.numeric(High.Grade), x='')) 
box1 = box1 + geom_boxplot() + coord_flip() 

box2 = box1 + scale_y_continuous(labels=ordLabels,breaks=1:15)

box3=box2 + labs(title=boxplot_title,
                     x =NULL, 
                     y = 'Grade Level',
                     caption = source.text)

box4 = box3 + theme(panel.background = element_rect(fill = "white",
                                                    colour = "grey50"),
                    plot.title = element_text(hjust = 0.5),
                    plot.caption = element_text(hjust=0))
box4
```

####Exercise 4 - Histogram
Now I make a histogram for the variable FREE LUNCH from the washington education data. First I create a data frame for ggplot to plot using only complete cases
```{r}
eduwa3=eduwa[complete.cases(eduwa$Free.Lunch),] #getting rid of missings
```

Misc elements: title
```{r}
hist.title = "Histogram of Number of Students Served Free Lunch in Washington Schools"
```
<a id='ex4'></a>
Create Plot
```{r}
base= ggplot(eduwa3,aes(x = Free.Lunch))  
h1= base + geom_histogram(color="black", fill="white", binwidth = 75,boundary=0) #changing width
h2= h1 + stat_bin(binwidth = 75, aes(label=..count..), 
                  geom = "text",boundary = 0,vjust=-0.5)

h3 = h2 + theme(panel.background = element_rect(fill = "white",
                                                    colour = "grey50"),
                    plot.title = element_text(hjust = 0.5),
                    plot.caption = element_text(hjust=0)) + labs(title=hist.title,
                     x ="Number of Students Served Free Lunch", 
                     y = "Frequency",
                     caption = source.text)

h3
```

####Exercise 5 - Histogram and density plot of student teacher ratio
First I create the data frame with the necessary variables (mean student teacher ratio by county)
```{r}
eduwa4=eduwa[complete.cases(eduwa$Student.Teacher.Ratio),]

eduwa5=as.data.frame(tapply(eduwa4$Student.Teacher.Ratio, 
       eduwa4$County, 
       mean))
names(eduwa5)=c('Student.Teacher.Ratio')

```
Misc elements:title
```{r}
hist.title2 = "Histogram of Average Student Teacher Ratios in Washington Counties"
```

<a id='ex5_1'></a>
Create Plot (Histogram)
```{r}
base = ggplot(eduwa5, aes(x=Student.Teacher.Ratio)) 

hist1 = base +  geom_histogram(color="black", fill="white", binwidth = 2 ,boundary=0)

hist2 = hist1 + theme(panel.background = element_rect(fill = "white",
                                                    colour = "grey50"),
                    plot.title = element_text(hjust = 0.5),
                    plot.caption = element_text(hjust=0)) + labs(title=hist.title2,
                     x ="Student Teacher Ratio", 
                     y = "Frequency",
                     caption = source.text)
hist3 = hist2 + scale_x_continuous(breaks=seq(0,24,2)) #adding more tick marks on the x axis

hist4 = hist3 + scale_y_continuous(breaks=seq(0,20,5)) # tick marks on y axis

hist5 = hist4 + geom_vline(aes(xintercept = mean(eduwa5$Student.Teacher.Ratio),colour="mean"),
                           show.legend  = TRUE) +
                geom_vline(aes(xintercept = median(eduwa5$Student.Teacher.Ratio),colour="median"),
                           show.legend  = TRUE) + 
                scale_color_manual(name = NULL, 
                                   values = c(median = "blue", mean = "red"))
hist5
```
Misc elements: title
```{r}
density.title = "Density Plot of Average Student Teacher Ratios in Washington Counties"
```

<a id='ex5_2'></a>
Create Plot (Density Plot)
```{r}
den1 = base + geom_density(color="black", fill="white")

den2 = den1 + theme(panel.background = element_rect(fill = "white",
                                                    colour = "grey50"),
                    plot.title = element_text(hjust = 0.5),
                    plot.caption = element_text(hjust=0)) + labs(title=density.title,
                     x ="Student Teacher Ratio", 
                     y = "Density",
                     caption = source.text)

den3 = den2 + scale_x_continuous(breaks=seq(0,24,2))

den4 = den3 + geom_vline(aes(xintercept = mean(eduwa5$Student.Teacher.Ratio),colour="mean"),
                           show.legend  = TRUE) +
                geom_vline(aes(xintercept = median(eduwa5$Student.Teacher.Ratio),colour="median"),
                           show.legend  = TRUE) + 
                scale_color_manual(name = NULL, 
                                   values = c(median = "blue", mean = "red"))

den4

```

